---
title: "Data Preparation"
author: "Md Zia Uddin Foisal"
date: "2023-10-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Set the Working Directory
setwd("E:/Bangladesh-CCDR-World-Bank/")

#Remove everything that is in the workspace
rm(list=ls())
```


```{r, echo=FALSE, warning=FALSE}
#Utils
library(readr)
library(dplyr)
library(tidyr)
library(purrr)
library(sf)
library(stringr)
library(glue)
library(janitor)
library(rio)
library(readxl)
# remotes::install_github("Ebukin/devPTIpack")
library(devPTIpack)
library(stringr)
```

##Read the administrative boundaries

```{r}
#District and Upazila BBS

bgd_adm0 <- st_read("./CCDR_Dashboard/raw_data/BGD_ADM23.gpkg", layer = "BGD_ADM0") %>% 
  clean_names()

division_shp <- st_read("./CCDR_Dashboard/raw_data/BGD_ADM23.gpkg", layer = "BGD_ADM1")%>%
  rename(division = ADM1_NAME) %>% 
  clean_names()

district_shp <- st_read("./CCDR_Dashboard/raw_data/BGD_ADM23.gpkg", layer = "BGD_ADM2") %>%
  rename(division = ADM1_NAME, district = ADM2_NAME) %>% 
  clean_names()
  
upazila_shp <- st_read("./CCDR_Dashboard/raw_data/BGD_ADM23.gpkg", layer = "BGD_ADM3") %>% 
  rename(division = ADM1_NAME , district = ADM2_NAME, upazila = ADM3_NAME) %>% 
  clean_names()

```

##Preparing the Diaster and Climate Risk Data

```{r}
#Reading Data
#Natural Hazard Data
natural_hazards <- rio::import_list("./CCDR_Dashboard/raw_data/BGD_RSK_23.xlsx")

hazards_district <- natural_hazards[["ADM2_summary"]] %>% 
  as_tibble()%>% 
  rename(division = ADM1_NAME, district = ADM2_NAME) %>% 
  #select(-ADM0_CODE, -ADM0_NAME, -ADM1_CODE, -ADM2_CODE) %>% 
  rename(dist_population = ADM2_POP) %>% 
  mutate(dist_population = dist_population/1000) %>% 
  pivot_longer(dist_population:`DR_S1_50p_mean` ,names_to = "indicator", values_to = "value") %>% 
  mutate(polygon = "District") %>% 
  arrange(district) %>% 
  mutate(value = case_when(
    indicator = str_detect(indicator, "%") ~ value*100,
    TRUE ~ value
  ))%>% 
  clean_names()

hazards_upazila <- natural_hazards[["ADM3_summary"]]%>% 
  as_tibble() %>% 
  #select(-ADM0_CODE, -ADM1_CODE, -ADM2_CODE, -ADM3_CODE, -ADM0_NAME) %>% 
  rename(division = ADM1_NAME , district = ADM2_NAME, upazila = ADM3_NAME) %>%
  rename(upazila_poplation = ADM3_POP) %>% 
  mutate(upazila_poplation = upazila_poplation/1000) %>%
  pivot_longer(upazila_poplation:`DR_S1_50p_mean` ,names_to = "indicator", values_to = "value") %>% 
  mutate(polygon = "Upazila") %>% 
  arrange(upazila) %>% 
  mutate(value = case_when(
    indicator = str_detect(indicator, "%") ~ value*100,
    TRUE ~ value
  )) %>% 
  clean_names()

#Matching Upazila Names with official boundary
upazila_1 <- upazila_shp$upazila

# hazards_upazila <- hazards_upazila %>%
#   bind_cols(upazila1= upazila_1) %>% 
#   select(-upazila) %>% 
#   rename(upazila = upazila1)


hazards <- bind_rows(hazards_district, hazards_upazila) %>% 
  mutate(domain = "Natural Hazards")
```

##Bringing Relative Wealth Index


```{r}
##Bringing in Relative Wealth Indices at district and Upazila level

rwi_data <- import_list("/CCDR_Dashboard/raw_data/BGD_RWI_summary.xlsx")

#district rwi 
rwi_district <- rwi_data$ADM2_RWI_mean %>% 
  pivot_longer(rwi_mean, names_to = "indicator", values_to = "value") %>% 
  mutate(polygon = "District", domain = "Relative Wealth Index") %>% 
  arrange(district)
  
#Upazila rwi
rwi_upazila <- rwi_data$ADM3_RWI %>% 
  pivot_longer(rwi_mean, names_to = "indicator", values_to = "value") %>%
  mutate(polygon = "Upazila", domain = "Relative Wealth Index") %>% 
  arrange(upazila)

#rwi_combined
rwi_final <- 
  bind_rows(rwi_district, rwi_upazila)


legend <- natural_hazards[["Legend"]] %>% 
  as_tibble()
```


##Preparing the Socio-economic data


```{r}
# #Socio-economic Indicator

development_indicators <- readxl::read_excel("/CCDR_Dashboard/raw_data/socioeconomic_indicators_for_bgd_dash_1.xlsx", na="NA")

development_indicators$indicator_1 <- lapply(development_indicators$indicator_1 , str_trim)
development_indicators$polygon <- "District"


#since Division is missing in development indicators, bringing in division
# prov <- district_shp %>% 
#   as_tibble() %>% 
#   select(division, district)

development_indicators <- development_indicators %>%
  mutate(domain = "Development Outcomes") %>%
  arrange(district)
```

##Combinning all the dataset

```{r}
#Combining all data
data <- hazards %>% 
  bind_rows(development_indicators) %>% 
  bind_rows(rwi_final) %>% 
  mutate(value = 
           case_when(
          domain == "Development Outcomes" ~ round(value, 2),
          domain == "Natural Hazards" ~ round(value, 2),
          domain == "Relative Wealth Index" ~ round(value, 2)
          )
  )
```


```{r}
unique_names_data <- data %>%
  distinct(indicator, indicator_1, unit,	source)

sdg_meta <- readxl::read_excel("/CCDR_Dashboard/raw_data/sdg_plus_metadata_glossary.xlsx")

unique_names_data$definition <- sdg_meta$Definition[match(unique_names_data$indicator, sdg_meta$Indicators_short_name)]

#write.csv(unique_names_data, "./CCDR_Dashboard/raw_data/legend_bgd_data_3.csv", row.names = FALSE)
```


```{r}
#Reading in Color mapping file for data Join
legend_data <- readxl::read_excel("./CCDR_Dashboard/raw_data/legend_bgd_data.xlsx", na= " ") %>%
  mutate(unit = replace_na(unit, ""))


data$unit <- legend_data$unit[match(data$indicator, legend_data$indicator)]
data$indicator_1 <- legend_data$indicator_1[match(data$indicator, legend_data$indicator)]
data$source <- legend_data$source[match(data$indicator, legend_data$indicator)]
data$context <- legend_data$context[match(data$indicator, legend_data$indicator)]
data$definition <- legend_data$definition[match(data$indicator, legend_data$indicator)]


data %>%
  mutate(value = case_when(
    indicator == "Total population in the division (in thousands)" ~ round(value, 0),
    indicator == "Total population in the district (in thousands)" ~ round(value, 0),
    indicator == "Total population in the upazila (in thousands)" ~ round(value, 0),
    indicator == "Expected annual mortality due to river floods (population count)" ~ round(value, 0),
    indicator == "Expected annual mortality due to coastal floods (population count)" ~ round(value, 0),
    indicator == "Expected annual exposure of population to extreme heat stress hazard (population count)" ~ round(value, 0)
    ))
    # indicator == "Expected annual exposure of population to very strong heat stress hazard (population count)" ~ round(value, 0),
    # indicator == "Population exposed to landslide hazard (population count)" ~ round(value, 0), TRUE ~ value ))


write_rds(data, file = "E:/RDL - Bangladesh Integrated Spatial Database/data_prep_bisd/Final RDS V1/data.RDS")
```

#Check the duplicates in the dataset

```{r}
duplicate <- {data} %>%
    dplyr::group_by(district, upazila, indicator_1) %>%
    dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
    dplyr::filter(n > 1L)
```



##Writing the Bangladesh Boundary data in RDS format

```{r}
bgd_shp <- list(District = district_shp, Upazila = upazila_shp)
bgd_shp %>% write_rds("/CCDR_Dashboard/data/bgd_shp.RDS")

```

```{r}
# bgd_pca <- data %>%
#   filter(polygon == "District")%>%
#   pivot_wider(id_cols = c("district"),
#     names_from = indicator_1,
#               values_from = value)

```

bgd_pca <- data %>%
  pivot_wider(names_from = indicator_1,
              values_from = value,
              id_cols = c("district", "upazila"),
              values_fn = {mean})

```{r}
library(sf)

bgd_adm2 <- st_read("./CCDR_Dashboard/raw_data/BGD_ADM23.gpkg", layer = "BGD_ADM2")


# bgd_indicators <- left_join(bgd_adm2, bgd_pca, by = c("adm2_en" = "district")) %>%
#   mutate(admin2Pcod = paste0("Bangladesh", bgd_adm2$adm2_pcode)) %>% 
#   mutate(admin0Pcod = bgd_adm2$adm0_en) %>%
#   mutate(admin2Name = bgd_adm2$adm2_en)

#write_xlsx(bgd_indicators, "E:\\RDL - Bangladesh Integrated Spatial Database\\data_prep_bisd\\bgd_indicators.xlsx")

```


```{r}
bgd_shp_district_PTI <- bgd_adm2 %>%
  mutate(admin2Pcod = paste0("Bangladesh", bgd_adm2$adm2_pcode)) %>% 
  mutate(admin0Pcod = bgd_adm2$adm0_en) %>%
  mutate(admin2Name = bgd_adm2$adm2_en)%>%
  select(admin0Pcod, admin2Pcod, admin2Name, geom)


#Making a List
bgd_shp <- list(admin2_District = bgd_shp_district_PTI)

bgd_shp <- 
  bgd_shp %>% 
  map(~{.x %>% st_transform("+proj=longlat +datum=WGS84")})

## https://stackoverflow.com/questions/68478179/how-to-resolve-spherical-geometry-failures-when-joining-spatial-data
sf_use_s2(FALSE)

bgd_shp$admin0_Country <-
  bgd_shp$admin2_District %>% 
  group_by(admin0Pcod) %>% 
  summarise() %>% 
  mutate(admin0Name = "Bangladesh") %>% 
  st_make_valid()


pti_shps <- NULL
pti_shps$admin0_Country  <- bgd_shp$admin0_Country 
pti_shps$admin2_District <- bgd_shp$admin2_District


pti_shps %>% 
  write_rds("E:/RDL - Bangladesh Integrated Spatial Database/data_prep_bisd/Final RDS V1/bgd_geometries.RDS", compress = "gz")

pti_shps$admin0_Country %>% st_crs(4326)
pti_shps$admin1_District %>% st_crs(4326)

#validate_geometries(pti_shps)
```





```{r}
pivoted_df <- development_indicators %>%
  select(division, district, indicator_1, value ) %>% 
  pivot_wider(names_from = indicator_1,
              values_from = value)

bgd_indicators <- left_join(bgd_adm2, pivoted_df, by = c("adm2_en" = "district")) %>%
  mutate(admin2Pcod = paste0("Bangladesh", bgd_adm2$adm2_pcode)) %>% 
  mutate(admin0Pcod = bgd_adm2$adm0_en) %>%
 mutate(admin2Name = bgd_adm2$adm2_en)

write_xlsx(bgd_indicators, "E:\\RDL - Bangladesh Integrated Spatial Database\\data_prep_bisd\\Data Updated 24October\\bgd_indicators.xlsx")
```



##Writing climate meta data
```{r}
###Metadata for PTI
metadata_climate <- import_list("E:/RDL - Bangladesh Integrated Spatial Database/data_prep_bisd/Data Updated 24October/Bgd_meta_data_climate.xlsx")
#devPTIpack::validate_metadata(metadata_climate)

metadata_climate %>% 
  write_rds("E:/RDL - Bangladesh Integrated Spatial Database/data_prep_bisd/Final RDS V1/bgd_metadata_climate.RDS")
```


##PTI shapfile Preparation

```{r}
# #PTI Shapefile preparation
# 
# nums1 <- seq(1:8)
# nums1 <- paste0(0, nums1)
# 
# nums2 <- as.character(seq(10,64)) 
# 
# #Concatenating together
# nums <- c(nums1, nums2)
# 
# bgd_shp_district_PTI <- st_read("data/shapefile_district/bgd_indicators.shp") %>% 
#   st_as_sf() %>% 
#   clean_names() %>% 
#   select(division, district, geometry) %>% 
#   arrange(district) %>% 
#   rename(admin1Name = division)  %>% 
#   mutate(admin1Pcod = paste0("Bangladesh", nums)) %>% 
#   mutate(admin0Pcod = "Bangladesh") %>% 
#   select(admin0Pcod, admin1Pcod, admin1Name, geometry)
# 
# #Making a List
# bgd_shp <- list(admin2_District = bgd_shp_district_PTI)
# 
# bgd_shp <- 
#   bgd_shp %>% 
#   map(~{.x %>% st_transform("+proj=longlat +datum=WGS84")})
# 
# ## https://stackoverflow.com/questions/68478179/how-to-resolve-spherical-geometry-failures-when-joining-spatial-data
# sf_use_s2(FALSE)
# 
# bgd_shp$admin0_Country <-
#   bgd_shp$admin2_District %>% 
#   group_by(admin0Pcod) %>% 
#   summarise() %>% 
#   mutate(admin0Name = "Bangladesh") %>% 
#   st_make_valid()
# 
# 
# pti_shps <- NULL
# pti_shps$admin0_Country  <- bgd_shp$admin0_Country 
# pti_shps$admin2_District <- bgd_shp$admin2_District
# 
# 
# pti_shps %>% write_rds("./CCDR_Dashboard/data/bgd_geometries.rds", compress = "gz")
# 
# pti_shps$admin0_Country %>% st_crs()
# pti_shps$admin2_District %>% st_crs()
# 
# validate_geometries(pti_shps)
```

##Preparing data for PCA

```{r}

#For pca (Using the data for PTI in wide format)
#Using data in original format 
metadata_climate <- import_list("./CCDR_Dashboard/raw_data/Bgd_meta_data_climate.xlsx")

data_pca <- metadata_climate$admin2_District
data_pca %>% 
  write_rds("./CCDR_Dashboard/data/data_pca.RDS")

```




```{r}
legend_rds <- metadata_climate[["metadata"]]
write_rds(legend_rds, "./CCDR_Dashboard/data/legend.RDS")
```


```{r}
district <- read.csv("./CCDR_Dashboard/raw_data/district.csv")
write_rds(legend_rds, "./CCDR_Dashboard/data/district.RDS")
```
